\documentclass{sig-alternate-2013}

\newfont{\mycrnotice}{ptmr8t at 7pt}
\newfont{\myconfname}{ptmri8t at 7pt}
\let\crnotice\mycrnotice%
\let\confname\myconfname%

\permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from permissions@acm.org.}
\conferenceinfo{ITCS'14,}{January 12--14, 2014, Princeton, New Jersey, USA.}
\copyrightetc{Copyright 2014 ACM \the\acmcopyr}
\crdata{978-1-4503-2698-8/14/01\ ...\$15.00.\\
http://dx.doi.org/10.1145/2554797.2554824}

\clubpenalty=10000
\widowpenalty = 10000

\usepackage{amsmath,amssymb}
\usepackage[usenames]{color}
\usepackage{graphicx}
\usepackage{subfigure}
%%\usepackage[round,colon]{natbib}

\input{latex_macros}

\begin{document}
\title{Attribute-Efficient Evolvability of Linear Functions} 
\numberofauthors{2}
\author{
\alignauthor
Elaine Angelino \titlenote{This author is supported in part by a grant from the
National Library of Medicine (4R01LM010213-04) and NSF grant CCF-09-64401.} \\
	\affaddr{Harvard University}\\
	\email{elaine@eecs.harvard.edu}
\alignauthor
Varun Kanade\titlenote{This author is supported by a Simons Postdoctoral
Fellowship.}\\
	\affaddr{University of California, Berkeley}\\
	\email{vkanade@eecs.berkeley.edu}
}

\maketitle

\begin{abstract}
\input{files/abstract}
\end{abstract}

\category{F.2}{Analysis of Algorithms and Problem Complexity}{General}
%\category{Machine Learning Theory}

\keywords{Evolvability, Computational Learning Theory, Regression}

\section{Introduction}
\label{sec:introduction}
\input{files/introduction}

\section{Model and Preliminaries}
\label{sec:notation}
\input{files/notation}

\section{Evolving Sparse Linear Functions}
\label{sec:algorithms}
\input{files/sparse_linear}

\subsection{Evolving Sparse Linear Functions Using $\optsel$}
\label{sec:greedy}
\input{files/greedy}

% \section{Evolving Sparse LTFs}
% \label{sec:sparse_LTF}
% \input{files/sparse_LTF}
% 
% \section{Evolving Sparse Disjunctions}
% \label{sec:sparse_disj}
% \input{files/sparse_disj}

\section{Conclusion and Future Work}
\label{sec:conclusion}
\input{files/conclusion.tex}

\subsection*{Acknowledgments} 
We would like to thank Leslie Valiant for helpful discussions and comments on an
earlier version of this paper. We are grateful to Frank Solomon for discussing
biological aspects related to this work.

% \bibliography{all-refs}
% \bibliographystyle{plain}
\input{files/all-refs.tex}

\newpage
\appendix 
\section{Omitted Proofs}

\subsection{Proofs from Section~\ref{sec:notation-class}}
\label{app:notation-class}
\input{files/app_notation}

\subsection{Proofs from Section~\ref{sec:sparse_linear}}
\label{app:sparse_linear}
\input{files/app_sparse_linear}

\subsection{Proofs from Section~\ref{sec:greedy}}
\label{app:greedy}
\input{files/app_greedy}


\end{document}
