\begin{thebibliography}{10}

\bibitem{Alon:2006}
Uri Alon.
\newblock {\em An Introduction to Systems Biology: Design Principles of
  Biological Circuits}.
\newblock Chapman and Hall/CRC, Boca Raton, FL, 2006.

\bibitem{Bach:2012}
Francis~R. Bach, Rodolphe Jenatton, Julien Mairal, and Guillaume Obozinski.
\newblock Optimization with sparsity-inducing penalties.
\newblock {\em Foundations and Trends in Machine Learning}, 4(1):1--106, 2012.

\bibitem{Balaji:2006}
S.~Balaji, Madan~M. Babu, Lakshminarayan~M. Iyer, Nicholas~M. Luscombe, and
  L.~Aravind.
\newblock Comprehensive analysis of combinatorial regulation using the
  transcriptional regulatory network of yeast.
\newblock {\em Journal of Molecular Biology}, 360(1):213--227, 2006.

\bibitem{Donoho:2009-sparse}
Alfred~M. Bruckstein, David~L. Donoho, and Michael Elad.
\newblock From sparse solutions of systems of equations to sparse modeling of
  signals and images.
\newblock {\em SIAM Review}, 51(1):34--81, 2009.

\bibitem{Candes:2008}
E.~J. Candes and M.~B. Wakin.
\newblock An introduction to compressive sampling.
\newblock {\em IEEE Signal Processing Magazine}, 25(2):21--30, 2008.

\bibitem{Daniel:1999}
Cuthbert Daniel and Fred~S. Wood.
\newblock {\em Fitting Equations to Data: Computer Analysis of Multifactor
  Data}.
\newblock John Wiley \& Sons, Inc., New York, NY, USA, 2nd edition, 1999.

\bibitem{Donoho:2006-recovery}
David~L. Donoho, Michael Elad, and Vladimir~N. Temlyakov.
\newblock Stable recovery of sparse overcomplete representations in the
  presence of noise.
\newblock {\em IEEE Transactions on Information Theory}, 52(1):6--18, 2006.

\bibitem{Elad:2010}
M.~Elad.
\newblock {\em Sparse and Redundant Representations: From Theory to
  Applications in Signal and Image Processing}.
\newblock Springer, 2010.

\bibitem{Feldman:2008-evolvability}
Vitaly Feldman.
\newblock Evolution from learning algorithms.
\newblock In {\em Proceedings of the Symposium on the Theory of Computation
  (STOC)}, 2008.

\bibitem{Feldman:2009-robustness}
Vitaly Feldman.
\newblock Robustness of evolvability.
\newblock In {\em Proceedings of the Conference on Learning Theory (COLT)},
  2009.

\bibitem{Feldman:2011-LTF}
Vitaly Feldman.
\newblock Distribution-independent evolution of linear threshold functions.
\newblock In {\em Proceedings of the Conference on Learning Theory (COLT)},
  2011.

\bibitem{Hartwell:1999}
L.~H. Hartwell, J.~J. Hopfield, S.~Leibler, and A.~W. Murray.
\newblock From molecular to modular cell biology.
\newblock {\em Nature}, 402(6761 Suppl):C47--C52, 1999.

\bibitem{Hastie:2001}
Trevor Hastie, Robert Tibshirani, and Jerome Friedman.
\newblock {\em The Elements of Statistical Learning}.
\newblock Springer New York Inc., New York, NY, USA, 2001.

\bibitem{Barabasi:2000}
H.~Jeong, B.~Tombor, R.~Albert, Z.~N. Oltvai, and A.~L. Barabasi.
\newblock The large-scale organization of metabolic networks.
\newblock {\em Nature}, 407(6804):651--654, 2000.

\bibitem{KST:2009}
Adam~Tauman Kalai, Alex Samorodnitsky, and Shang-Hua Teng.
\newblock Learning and smoothed analysis.
\newblock In {\em Proceedings of IEEE Conference on the Foundations of Computer
  Science (FOCS)}, 2009.

\bibitem{Kanade:2012-thesis}
Varun Kanade.
\newblock {\em Computational Questions in Evolution}.
\newblock PhD thesis, Harvard University, 2012.

\bibitem{KVV:2010-drift}
Varun Kanade, Jennifer~Wortman Vaughan, and Leslie~G. Valiant.
\newblock Evolution with drifting targets.
\newblock In {\em Proceedings of the Conference on Learning Theory (COLT)},
  2010.

\bibitem{KKM:2013}
Daniel~M. Kane, Adam Klivans, and Raghu Meka.
\newblock Learning halfspaces under log-concave densities: Polynomial
  approximations and moment matching.
\newblock In {\em Proceedings of the Conference on Learning Theory (COLT)},
  2013.

\vfill\eject
\bibitem{Kearns:1998}
Michael Kearns.
\newblock Efficient noise-tolerant learning from statistical queries.
\newblock {\em J. ACM}, 45(6):983--1006, 1998.

\bibitem{Littlestone:1988}
Nick Littlestone.
\newblock Learning quickly when irrelevant attributes abound: A new
  linear-threshold algorithm.
\newblock {\em Mach. Learn.}, 2(4):285--318, 1988.

\bibitem{Luscombe:2004}
Nicholas~M. Luscombe, M.~Madan~Babu, Haiyuan Yu, Michael Snyder, Sarah~A.
  Teichmann, and Mark Gerstein.
\newblock Genomic analysis of regulatory network dynamics reveals large
  topological changes.
\newblock {\em Nature}, (7006):308--312.

\bibitem{Mallat:2008}
Stephane Mallat.
\newblock {\em A Wavelet Tour of Signal Processing, Third Edition: The Sparse
  Way}.
\newblock Academic Press, 3rd edition, 2008.

\bibitem{Murray:2011}
Douglas~B. Murray, Ken Haynes, and Masaru Tomita.
\newblock Redox regulation in respiring \textit{Saccharomyces cerevisiae}.
\newblock {\em Biochimica et Biophysica Acta}, 1810(10):945--958, 2011.

\bibitem{Natarajan:1995}
B.~K. Natarajan.
\newblock Sparse approximate solutions to linear systems.
\newblock {\em SIAM J. Comput.}, 24(2):227--234, 1995.

\bibitem{Olshausen:1997}
Bruno~A. Olshausen and David~J. Fieldt.
\newblock Sparse coding with an overcomplete basis set: A strategy employed by
  {V1}?
\newblock {\em Vision Research}, 37:3311--3325, 1997.

\bibitem{ST:2004}
Daniel~A. Spielman and Shang-Hua Teng.
\newblock Smoothed analysis of algorithms: Why the simplex algorithm usually
  takes polynomial time.
\newblock {\em Journal of the ACM}, 51(3), 2004.

\bibitem{Tropp:2004-greed}
Joel~A. Tropp.
\newblock Greed is good: Algorithmic results for sparse approximation.
\newblock {\em IEEE Transactions on Information Theory}, 50(10):2231--2242,
  2004.

\bibitem{Valiant:2009-evolvability}
Leslie~G. Valiant.
\newblock Evolvability.
\newblock {\em Journal of the ACM}, 56(1):3:1--3:21, 2009.

\bibitem{Valiant:2012-real}
Paul Valiant.
\newblock Evolvability of real-valued functions.
\newblock In {\em Proceedings of Innovations in Theoretical Computer Science
  (ITCS)}, 2012.

\bibitem{Wagner:2001}
Andreas Wagner and David~A. Fell.
\newblock The small world inside large metabolic networks.
\newblock {\em Proceedings of the Royal Society of London. Series B: Biological
  Sciences}, 268(1478):1803--1810, 2001.

\bibitem{Watts:1998}
Duncan~J. Watts and Steven~H. Strogatz.
\newblock Collective dynamics of `small-world' networks.
\newblock {\em Nature}, 393(6684):440--442, 1998.

\end{thebibliography}
