In this work, we provided simple evolutionary mechanisms for evolving sparse
linear functions, under a large class of distributions. These evolutionary
algorithms have the desirable properties that the representations used are
themselves sparse linear functions, and that they are attribute-efficient in the
sense that the number of generations required for evolution to succeed is
independent of the total number of attributes.

Strong negative results are known for distribution-independent evolvability of
boolean functions, \eg even the class of conjunctions is not
evolvable~\cite{Feldman:2011-LTF}. However, along the lines of this work, it is
interesting to study whether under restricted classes of distributions,
evolution is possible for simple concept classes, using representations of low
complexity. Currently, even under (biased) product distributions, no
evolutionary mechanism is known for the class of disjunctions, except via
Feldman's general reduction from CSQ algorithms. Even if the queries made by the
CSQ algorithm are simple, Feldman's reduction uses intermediate representations
that randomly combine queries made by the algorithm, making the representations
quite complex.

A natural extension of our current results would be to study fixed-degree sparse
polynomials. Another interesting direction is to study circuits with sigmoidal
or other non-linear filters on the gates, which arise naturally in molecular
systems.  The suitable class of boolean functions to study is low-weight
threshold functions, which includes disjunctions and conjunctions. The class of
smooth bounded distributions may be an appropriate starting place for studying
evolvability of these classes. For example, is the class of low-weight threshold
functions evolvable under smooth distributions, or at least log-concave
distributions? 
